---
title: "INTEGER PROGRAMMING ASSIGNMENT"
author: "SAI ROHITH GUDA"
output:
  word_document: default
  html_document: default
  pdf_document: default
date: "11-19-23"
---
AP is a package delivery service that assures overnight delivery in the continental United States.
The firm operates hubs in key cities and airports around the country. Packages are received at hubs before being delivered to intermediate hubs or their final destination.

The manager of the AP hub in Cleveland is concerned about labor expenses and is looking into alternatives.

identifying the most effective strategy to schedule staff. The hub is open seven days a week and the amount of parcels handled changes from day to day.

Minimum number of workers required each day

Sunday 18 
Monday 27 
Tuesday 22
Wednesday 26
Thursday 25
Friday 21 
Saturday 19

shifts constraints are 

sunday : There are two  off, shifts 1 and 7, and the limitation is:
0w1+1w2+ 1w3+ 1w4+ 1w5+ 1w6+ 0w7


Monday : There are two  off's, shifts 1 and 2, and the limitation is
0w1+0w2+ 1w3+ 1w4+ 1w5+ 1w6+ 1w7

Tuesday :There are two  off's, shifts 2 and 3, and the limitation is
1w1+0w2+ 0w3+ 1w4+ 1w5+ 1w6+ 1w7

Wednesday: There are two  off's, shifts 3 and 4, and the limitation is
1w1+ 1w2+ 0w3+ 0w4+ 1w5+ 1w6+ 1w7

Thursday : There are two  off's, shifts 4 and 5, and the limitation is
1w1+ 1w2+ 1w3+ 0w4+ 0w5+ 1w6+ 1w7

Friday : There are two  off's, shifts 5 and 6, and the limitation is
1w1+1w2+ 1w3+ 1w4+ 0w5+ 0w6+ 1w7

Saturday : There are two  off's, shifts 6 and 7, and the limitation is:
1w1+1w2+ 1w3+ 1w4+ 1w5+ 0w6+ 0w7


```{r}
# Install and load the lpSolve package
if (!require(lpSolve)) {
    install.packages("lpSolve")
    library(lpSolve)
}

# Define the coefficients of the objective function (total wages for each shift)
objective_coefficients <- c(775, 800, 800, 800, 800, 775, 750)

# Define the constraints matrix
# Each row corresponds to a day, and each column corresponds to a shift
constraints <- matrix(c(
    0, 1, 1, 1, 1, 1, 0, # Sunday
    0, 0, 1, 1, 1, 1, 1, # Monday
    1, 0, 0, 1, 1, 1, 1, # Tuesday
    1, 1, 0, 0, 1, 1, 1, # Wednesday
    1, 1, 1, 0, 0, 1, 1, # Thursday
    1, 1, 1, 1, 0, 0, 1, # Friday
    1, 1, 1, 1, 1, 0, 0  # Saturday
), nrow = 7, byrow = TRUE)

# Define the right-hand side of the constraints (workers required each day)
constraints_rhs <- c(18, 27, 22, 26, 25, 21, 19)

# Define the constraints direction (greater than or equal to)
constraints_dir <- rep(">=", 7)

# Define the integer constraint for each decision variable
int_constraints <- rep(1, 7)

# Solve the linear programming problem
solution <- lp("min", objective_coefficients, constraints, constraints_dir, constraints_rhs, all.int = TRUE)

# Output the solution
if (solution$status == 0) {
    print("Optimal solution found:")
    print(solution$solution)
    
    cat("Total wage cost:", sum(solution$solution * objective_coefficients), "\n")
} else {
    print("Solution not found")
}
## substituting in the decision variables

```
```{r}

 
    
constraints <- matrix(c(
    0, 4, 5, 0, 8, 1, 0, # Sunday
    0, 0, 5, 0, 8, 1, 13, # Monday
    2, 0, 0, 0, 8, 1, 13, # Tuesday
    2, 4, 0, 0, 8, 1, 13, # Wednesday
    2, 4, 5, 0, 0, 1, 13, # Thursday
    2, 4, 5, 0, 0, 0, 13, # Friday
    2, 4, 5, 0, 8, 0, 0  # Saturday
), nrow = 7, byrow = TRUE)

 row.names(constraints) <- c("sunday","monday","tuesday", "wednesday",
                                "thursday", "friday", "saturday" )
 colnames(constraints)<- c("Shift 1", "Shift 2", "Shift 3", "Shift 4", "Shift 5", "Shift 6", "Shift 7")

    
print(constraints)
```



```{r}

    row.names(constraints) <- c("sunday","monday","tuesday", "wednesday",
                                "thursday", "friday", "saturday" )
    
 rowSums(constraints)
 #workers available for each day
 # total wage expense is 25675

```
